Conceptual and methodological problems with bias detection and avoidance in natural language processing

### 2021-06-22 thesis/02-cosine.Rmd
fixing small things

One should note that cosine distance is not exactly a distance measure, as it does not meet triangle inequality requirements. The triangle inequality formula says that for any triangle, the sum of the lengths of any two sides must be greater than or equal to the length of the remaining side. As shown in [discussion in stats.stackexchange.com](https://stats.stackexchange.com/questions/198080/proving-that-cosine-distance-function-defined-by-cosine-similarity-between-two-u)^[https://stats.stackexchange.com/questions/198080/proving-that-cosine-distance-function-defined-by-cosine-similarity-between-two-u] in the case of cosine distance it would have to fulfill this equation 1+cos-sim(𝐴,𝐶) < cos-sim(𝐴,𝐵) + cos-sim(𝐵,𝐶). If one chooses specific unit vectors it is easy to demonstrate that the triangle inequality is not preserved.

@@

\textbf{"jew"       : ["judaism", "jew", "synagogue", "torah", "rabbi"]}

@@

\textbf{"muslim"    : ["islam", "muslim", "mosque", "quran", "imam"]}

@@

\textbf{"jew"       : ["greedy", "cheap", "hairy", "liberal"]}

@@

\textbf{"muslim"       : ["violent", "terrorist", "dirty", "uneducated"]}

@@

\textbf{["judaism", "christianity", "islam"]}

@@

\textbf{["synagogue", "church", "mosque"]}

@@

\newpage

>>>>

One should note that cosine distance is not exactly a distance measure, as it does not meet triangle inequality requirements. The triangle inequality formula says that for any triangle, the sum of the lengths of any two sides must be greater than or equal to the length of the remaining side. As shown in [discussion in stats.stackexchange.com](https://stats.stackexchange.com/questions/198080/proving-that-cosine-distance-function-defined-by-cosine-similarity-between-two-u)^[https://stats.stackexchange.com/questions/198080/proving-that-cosine-distance-function-defined-by-cosine-similarity-between-two-u] in the case of cosine distance it would have to fulfill this equation $1+\mathsf{cos-sim}(A,C) < \mathsf{cos-sim}(A,B) + \mathsf{cos-sim}(B,C)$. If one chooses specific unit vectors it is easy to demonstrate that the triangle inequality is not preserved.

@@

\begin{minipage}{1.0\textwidth}
\textbf{"jew"       : ["judaism", "jew", "synagogue", "torah", "rabbi"]}

@@

\textbf{"muslim"    : ["islam", "muslim", "mosque", "quran", "imam"]}
\end{minipage}

@@

\begin{minipage}{1.0\textwidth}
\textbf{"jew"       : ["greedy", "cheap", "hairy", "liberal"]}

@@

\textbf{"muslim"       : ["violent", "terrorist", "dirty", "uneducated"]}
\end{minipage}

@@

\begin{minipage}{1.0\textwidth}
\textbf{["judaism", "christianity", "islam"]}

@@

\textbf{["synagogue", "church", "mosque"]}
\end{minipage}

@@

### 2021-06-22 thesis/03-bayesian.Rmd
fixing small things

\vspace{1mm}
\footnotesize
```{r tableMuslimActive,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "100%"}
library(tidyverse)
muslimClass <- muslim %>% filter(protectedWord == "muslim")
neutralSample <- sample_n(filter(muslimClass,connection == "none"), 5)
humanSample <- sample_n(filter(muslimClass,connection == "human"), 5)
muslimVis <- muslimClass %>% filter(connection != "none" & connection !="human")
muslimVis <- rbind(muslimVis,neutralSample,humanSample)

@@

\vspace{1mm}
\footnotesize
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
library(tidyverse)
priestClass <- religion %>% filter(protectedWord == "priest")
neutralSample <- sample_n(filter(priestClass,connection == "none"), 5)
humanSample <- sample_n(filter(priestClass,connection == "human"), 5)
priestVis <- priestClass %>% filter(connection != "none" & connection !="human")
priestVis <- rbind(priestVis,neutralSample,humanSample)

>>>>

\vspace{1mm}
\footnotesize
```{r tableMuslimActive,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "100%", warning = FALSE, message = FALSE}
library(tidyverse)
muslimClass <- muslim %>% filter(protectedWord == "muslim")
neutralSample <- sample_n(filter(muslimClass,connection == "none"), 5)
humanSample <- sample_n(filter(muslimClass,connection == "human"), 5)
muslimVis <- muslimClass %>% filter(connection != "none" & connection !="human")
muslimVis <- rbind(muslimVis,neutralSample,humanSample)

@@

\vspace{1mm}
\footnotesize
```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", warning = FALSE, message = FALSE}
library(tidyverse)
priestClass <- religion %>% filter(protectedWord == "priest")
neutralSample <- sample_n(filter(priestClass,connection == "none"), 5)
humanSample <- sample_n(filter(priestClass,connection == "human"), 5)
priestVis <- priestClass %>% filter(connection != "none" & connection !="human")
priestVis <- rbind(priestVis,neutralSample,humanSample)

### 2021-06-22 thesis/_book/_main.tex
fixing small things

One should note that cosine distance is not exactly a distance measure,
as it does not meet triangle inequality requirements. The triangle
inequality formula says that for any triangle, the sum of the lengths of
any two sides must be greater than or equal to the length of the
remaining side. As shown in
\href{https://stats.stackexchange.com/questions/198080/proving-that-cosine-distance-function-defined-by-cosine-similarity-between-two-u}{discussion
in stats.stackexchange.com}\footnote{\url{https://stats.stackexchange.com/questions/198080/proving-that-cosine-distance-function-defined-by-cosine-similarity-between-two-u}}
in the case of cosine distance it would have to fulfill this equation
1+cos-sim(𝐴,𝐶) \textless{} cos-sim(𝐴,𝐵) + cos-sim(𝐵,𝐶). If one chooses
specific unit vectors it is easy to demonstrate that the triangle
inequality is not preserved.

>>>>

One should note that cosine distance is not exactly a distance measure,
as it does not meet triangle inequality requirements. The triangle
inequality formula says that for any triangle, the sum of the lengths of
any two sides must be greater than or equal to the length of the
remaining side. As shown in
\href{https://stats.stackexchange.com/questions/198080/proving-that-cosine-distance-function-defined-by-cosine-similarity-between-two-u}{discussion
in stats.stackexchange.com}\footnote{\url{https://stats.stackexchange.com/questions/198080/proving-that-cosine-distance-function-defined-by-cosine-similarity-between-two-u}}
in the case of cosine distance it would have to fulfill this equation
\(1+\mathsf{cos-sim}(A,C) < \mathsf{cos-sim}(A,B) + \mathsf{cos-sim}(B,C)\).
If one chooses specific unit vectors it is easy to demonstrate that the
triangle inequality is not preserved.
